---
title: "GP Redo of Powerpoint"
author: "cliffbeall"
date: '2023-05-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(vegan)
```

## NMDS Panels

These are derived from all baby NMDS plot, each panel shows one only the points fomr one time point and compares the NICU and home groups. A PERMANOVA was done at each time point. This is slide 4 in the "Neonate for Kyu.pptx" file.

```{r }
load("gp_counts_meta.Rdata")
set.seed(3658)
nmds_baby_all <- gp_2023_spec_counts[, -1] %>%
  filter(gp_2023_meta$Age == "Baby") %>%
  decostand(method = "total") %>%
  metaMDS(trymax = 200, autotransform = FALSE)
  baby_nmds_gg_df <- cbind(gp_2023_meta[gp_2023_meta$Age == "Baby", ],
                 scores(nmds_baby_all, display = "sites")
                 )


nmds_by_visit <- function(visit){
  baby_nmds_gg_df %>%
    filter(Visit == visit) %>%
      ggplot(aes(x = NMDS1, y = NMDS2, color = Location)) +
        geom_point() +
        coord_fixed(xlim = c(-1, 3), ylim = c(-2, 2.6)) +
        scale_color_manual(values = c("blue2", "red")) +
        labs(title = str_c("Visit ", visit, collapse = ""))
}
nmds_by_visit("1")
nmds_by_visit("2")
nmds_by_visit("3")
babyv1 <- gp_2023_meta$Age == "Baby" & gp_2023_meta$Visit == "1"
babyv2 <- gp_2023_meta$Age == "Baby" & gp_2023_meta$Visit == "2"
babyv3 <- gp_2023_meta$Age == "Baby" & gp_2023_meta$Visit == "3"

gp_2023_spec_frac <- decostand(gp_2023_spec_counts[, -1], method = "total")

set.seed(3658)
"PERMANOVA Visit 1"
adonis(gp_2023_spec_frac[babyv1, ] ~ Location, data = gp_2023_meta[babyv1,])
"PERMANOVA Visit 2"
adonis(gp_2023_spec_frac[babyv2, ] ~ Location, data = gp_2023_meta[babyv2,])
"PERMANOVA Visit 3"
adonis(gp_2023_spec_frac[babyv3, ] ~ Location, data = gp_2023_meta[babyv3,])
```

The PERMANOVA results are very similar to the previous, NMDSs don't look quite the same.

## Cesarean vs. Vaginal 3 month

This is to reproduce slide 7 of the powerpoint.

```{r}
set.seed(3658)
bb3mo <- gp_2023_meta$Age == "Baby" & gp_2023_meta$Visit == "3"
gp_3mo_baby_nmds <- gp_2023_spec_frac %>%
  filter(bb3mo) %>%
  as.matrix() %>%
  metaMDS(trymax = 200, autotransform = FALSE)
gp_3mo_baby_gg <- gp_2023_meta %>%
  filter(bb3mo) %>%
  cbind(scores(gp_3mo_baby_nmds, display = "sites"))
ggplot(gp_3mo_baby_gg, aes(x = NMDS1, y = NMDS2, color = Delivery)) +
  geom_point() +
  coord_fixed() +
  scale_color_manual(values = c("orange", "green4"))

adonis(gp_2023_spec_frac[bb3mo, ] ~ gp_2023_meta$Delivery[bb3mo])
```

## Breast vs. Formula Feeding

NMDS with PERMANOVA and Shannon. To reproduce Slide 8 of the powerpoint.

```{r}
ggplot(gp_3mo_baby_gg, aes(x = NMDS1, y = NMDS2, color = Breastfeeding)) +
  geom_point() +
  coord_fixed() +
  scale_color_manual(values = c("black", "hotpink", "purple", "brown"))
adonis(gp_2023_spec_frac[bb3mo, ] ~ gp_2023_meta$Breastfeeding[bb3mo])
```




